Thermodynamic fragments in magic state theories

### 2020-12-07 magic_paper.tex
Minor edits. Questions/tasks posed at end of unital fragment section.

\subsection{Magic in the unital fragment}\label{sec:unital}
The unital fragment encompasses the circuits which preserve the maximally mixed state ($\id/d$) and so it includes many important families of circuits.

@@

\subsection{Strange state distillation in the unital fragment}\label{sec:lcsu}

We would now like to study the constraints coming solely from majorization in the unital fragment. 
We consider the problem of purifying $\nn$ copies of a noisy strange state $\noisysn$ into a smaller number of copies $\tt$ of a less noisy strange state $\noisys(\epsilon')\too{\tt}$, with $\epsilon' \le \epsilon$ and $\tt \leq \nn$. 
Since the state $\frac{1}{3} \id$ is free, the interconversion problem is unaffected if we tensor in copies of it. 
This turns out to simplify the Lorenz curve analysis, and so we focus on the process,
\begin{equation}\label{eq:sudist}
	\noisys\too{\nn} \longrightarrow \noisys(\tt, \epsilon', \nn - \nn') \coloneqq  \noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{(\nn-\tt)},
\end{equation}
where all copies $\nn, \tt, \nn - \tt$ are even.
Since the initial state can be written in the form $\noisys(\nn, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\noisys}{\id/3}(x)$ for the generic noisy strange state $\noisys(\nn, \epsilon ,\tt)$.
This curve is defined at $9^\nn$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $2^3 = 8$ different forms, depending on whether the noise level $\epsilon$ is low/high (relative to $\frac{3}{7}$), the parity of the number of copies $\nn$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
The full details are provided in~\cref{app:lcsu_technical}.

Here we focus on the case of Lorenz curves of even copies $\nn, \tt$ and low noise levels ($\epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_\ii$, with associated multiplicities $\mlt_\ii$,
\begin{align}
	w_\ii &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2\ii}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{\nn-2\ii}, \\
	m_\ii &= 8^{2\ii}\binom{\nn}{2\ii},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
From this, it is readily seen that $\lc{\rho}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_\ii$, where
\begin{equation}
	x_\ii = \Phi_+ \left( 2\ii, \nn , \frac{8}{9} \right),
\end{equation}
with $\ii = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2\ii, \nn, p) \coloneqq \sum_{\dummy=0}^{\ii} \binom{\nn}{2\dummy} p ^{2\dummy} (1-p)^{\nn-2\dummy}.
\end{equation}
The value of the Lorenz curve at an elbow is given by
\begin{equation}
	\lc{\rho}{\id/3}(x_{\ii}) \equiv L_{\ii} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{\nn} \Phi_+\left(2i;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

@@

Lorenz curve comparison provides $9^\nn$ necessary constraints for the process~\cref{eq:sudist}.
\cref{thm:elbows} shows that they can be reduced to $\tt$ independent constraints.
However, in principle we can pick any location $x_\star$ at which the Lorenz curves are defined and we then get a necessary distillation constraint, 
\begin{equation}
	\lc{\noisys(\nn, \epsilon, 0)}{\frac{\id}{3}}(x_\star) \geq \lc{\noisys(\tt, \epsilon', \nn-\tt)}{\frac{\id}{3}}(x_\star).
\end{equation}

@@

The first elbow of the target state is located at $x'_0 = \frac{1}{9^{\tt}} > x_\star$, so in order to find the target state Lorenz curve coordinate $L'_\star$ at location $x_\star$, we can always interpolate between the origin and the first target state elbow, 
\begin{equation}
	\begin{split}
	L'_\star &= \frac{x_\star}{x'_0}L'_0 = \frac{1}{9^{\nn-\tt}}\left( \frac{5}{3} - \frac{8}{9}\epsilon'\ \right)^{\tt}\Phi_+\left(0;\tt,\frac{12-4\epsilon'}{15-8\epsilon'}\right) \\
	&= \frac{(3-4\epsilon')^{\tt}}{9^{\nn}}.
	\end{split}
\end{equation}
The process in~\cref{eq:sudist} is therefore possible if $L_\star \geq L'_\star$ or equivalently,
\begin{align}
	\frac{\nn}{\tt} &\geq \frac{\ln{(3-4\epsilon')}}{\ln{(3-4\epsilon)}}, \\
	\epsilon &\leq \frac{3}{4} - \frac{1}{4}(3-4\epsilon')^{\frac{\tt}{\nn}}.
\end{align}
As an example, the process with $(\nn, \tt) = (12, 4)$ and $\epsilon' = 0.05$ yields $\epsilon \leq 0.398$.
Numerical comparison of the curves with varying $\epsilon$ suggests that we require $\epsilon \leq 0.373$ for the process to be possible.
\nick{All numerical examples I have tested, for various different cases, are quite close}

\null\newpage
\null\newpage

>>>>

\section{Magic in the unital fragment}\label{sec:unital}
The unital fragment encompasses the circuits which preserve the maximally mixed state ($\id/d$) and so it includes many important families of circuits.

@@

\subsection{Distillation rate bounds for the unital fragment}\label{sec:lcsu}

We would now like to study the constraints coming solely from majorization in the unital fragment. 
We consider the problem of purifying $\nn$ copies of a noisy strange state $\noisysn$ into a smaller number of copies $\tt$ of a less noisy strange state $\noisys(\epsilon')\too{\tt}$, with $\epsilon' \le \epsilon$ and $\tt \leq \nn$. 
Since the state $\frac{1}{3} \id$ is free, the interconversion problem is unaffected if we tensor in copies of it. More precisely the associated Lorenz curve in invariant under tensoring in arbitrary many copies of this free state.
This turns out to simplify the Lorenz curve analysis, and so we focus on the process,
\begin{equation}\label{eq:sudist}
	\noisys\too{\nn} \longrightarrow \noisys(\tt, \epsilon', \nn - \nn') \coloneqq  \noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{(\nn-\tt)},
\end{equation}
where all copies $\nn, \tt, \nn - \tt$ are even.
Since the initial state can be written in the form $\noisys(\nn, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\noisys}{\id/3}(x)$ for the generic noisy strange state $\noisys(\nn, \epsilon ,\tt)$.
This curve is defined at $9^\nn$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $8$ different forms, depending on whether the noise level $\epsilon$ is greater than or less than $\frac{3}{7}$, the parity of the number of copies $\nn$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
The full details are provided in~\cref{app:lcsu_technical}.

Here we focus on the case of Lorenz curves of even copies $\nn, \tt$ and low noise levels ($\epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_\ii$, with associated multiplicities $\mlt_\ii$,
\begin{align}
	w_\ii &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2\ii}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{\nn-2\ii}, \\
	m_\ii &= 8^{2\ii}\binom{\nn}{2\ii},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
From this, it is readily seen that $\lc{\rho}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_\ii$, where
\begin{equation}
	x_\ii = \Phi_+ \left( 2\ii, \nn , \frac{8}{9} \right),
\end{equation}
with $\ii = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2\ii, \nn, p) \coloneqq \sum_{\dummy=0}^{\ii} \binom{\nn}{2\dummy} p ^{2\dummy} (1-p)^{\nn-2\dummy}.
\end{equation}
We provide technical details on the function $\Phi_+$ in the appendices.

The value of the Lorenz curve at an elbow is given by
\begin{equation}
	\lc{\rho}{\id/3}(x_{\ii}) \equiv L_{\ii} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{\nn} \Phi_+\left(2i;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

@@

Lorenz curve comparison provides $9^\nn$ necessary constraints for the process~\cref{eq:sudist}.
\cref{thm:elbows} shows that they can be reduced to $\tt$ independent constraints.
However, in principle we can pick any location $x_\star \in [0,1]$ at which the Lorenz curves are defined and we then get a necessary distillation constraint, 
\begin{equation}
	\lc{\noisys(\nn, \epsilon, 0)}{\frac{\id}{3}}(x_\star) \geq \lc{\noisys(\tt, \epsilon', \nn-\tt)}{\frac{\id}{3}}(x_\star).
\end{equation}

@@

The first elbow of the target state is located at $x'_0 = \frac{1}{9^{\tt}} > x_\star$, so in order to find the target state Lorenz curve coordinate $L'_\star$ at location $x_\star$, we can interpolate between the origin and the first target state elbow, 
\begin{equation}
	\begin{split}
	L'_\star &= \frac{x_\star}{x'_0}L'_0 = \frac{1}{9^{\nn-\tt}}\left( \frac{5}{3} - \frac{8}{9}\epsilon'\ \right)^{\tt}\Phi_+\left(0;\tt,\frac{12-4\epsilon'}{15-8\epsilon'}\right) \\
	&= \frac{(3-4\epsilon')^{\tt}}{9^{\nn}}.
	\end{split}
\end{equation}
The process in~\cref{eq:sudist} is therefore possible only if $L_\star \geq L'_\star$ or equivalently,
\begin{align}
	\frac{n'}{n} &\leq \frac{\ln{(3-4\epsilon)}}{\ln{(3-4\epsilon')}}, \\
	\epsilon &\leq \frac{3}{4} - \frac{1}{4}(3-4\epsilon')^{\frac{\tt}{\nn}}.
\end{align}
As an example, the process with $(\nn, \tt) = (12, 4)$ and $\epsilon' = 0.05$ yields $\epsilon \leq 0.398$.
Numerical comparison of the curves with varying $\epsilon$ suggests that we require $\epsilon \leq 0.373$ for the process to be possible.
\nick{All numerical examples I have tested, for various different cases, are quite close}

Moreover, for the problem of distilling pure magic states where $\epsilon'=0$ we obtain an upper bound on the unital fragment distillation rate $R=n/n'$ given by
\begin{equation}
R \le \frac{\ln (3-4 \epsilon)}{\ln 3},
\end{equation}
which holds for all finite $n$. In comparison, the bound from mana gives $ R \le $ \ddd{[GIVE VALUE]}.

\ddd{[This is good, and the bound coming from the first elbow seems very easy to obtain. Going forward:
\begin{enumerate}
\item Does the ``last elbow'' give the same bound? Is it as simple?
\item Numerically is this about as good as it gets from majorization?
\item How does this rate compare with existing literature?
\end{enumerate}
If this bound is better than existing literature then great. I suspect not though. However, remember the core message of our work is that we're ``incorporating physical restrictions'' into the magic analysis. Given this, one thing that would round the paper off could be if we re-played the ``first elbow'' bound for some non-trivial temperature. Meaning: choose a ``nice'' Hamiltonian and compute the first elbow constraint via the thermo-majorization distortion of the Wigner components. This would then give us something clearly new -- namely a bound with temperature dependence! People would find this interesting since nothing like it exists yet.}
\null\newpage
\null\newpage
